\textbf{ \Large{Reducing data traffic in wireless structural health monitoring systems using embedded computing}}

### 2015-02-09 Paper/PaperJahrSchlichSHM.tex
Abstract angepasst, UML-Diagram geteilt, kleine Änderungen am Literaturverzeichnis

Structural health monitoring systems
shm systems produce a lot of data
we reduce the data with fft
The dependability and the accuracy of structural health monitoring systems can be affected by sensor faults. 
In this paper, the design and implementation of a wireless structural health monitoring system, capable of decentralized autonomous fault detection, are presented. 
For self-detecting sensor faults, each sensor node predicts expected sensor data and compares it to the measured sensor data. 
In laboratory experiments, devised to validate the proposed approach, the data traffic is reduced.

@@

\linenumbers % Schaltet Zeilennummerierung ein
\modulolinenumbers[5] % nur jede 5. Zeile
\section*{Dictionary}

\begin{tabular}{|l|l|}
\hline 
Sensorknoten (SunSPOT) & sensor node \\ 
\hline 
einzelner Messsensor (Thermometer) & sensor \\ 
\hline 
Knoten im neuronalen Netz & neuron \\ 
\hline 
eine abgeschlossene Messungreihe & test run \\ 
\hline 
gemessene Werte & sensor data \\ 
\hline 
vorhergesagte Werte & predicted data \\ 
\hline 
durch Vorhersage erwartete Werte & expected data \\ 
\hline 
einzelner Messwert & measurement \\ 
\hline 
Test & laboratory experiments\\ 
\hline
tatsächliche, nicht virtuelle Messung & actual measurement \\
\hline
Messaufbau & test setup \\
\hline

\end{tabular}

@@

\todo{citation BILD!}
\begin{figure}[ht]
    \centering
    \includegraphics[scale=0.5]{figures/eDemoboard.png}
    \caption{Layout of the Sun SPOT sensor board}
    \label{fig:SPOT}
\end{figure}

@@

\autoref{fig:UML} describes the classes of the \texttt{sensornode} package.
The package \texttt{sensornode} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, which are embedded directly into the sensor nodes.
The \texttt{AccelerationSampler} class is responsible for measuring the acceleration.
There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
Once the acceleration exceeds a threshold, the \colorbox{cyan}{second phase is entered} by increasing the sampling rate. 
The measured values are stored into an array.
The different phases are indicatad by lighting different LEDs.
The \texttt{FFT} class performs a fast Fourier transform on the measured accelerations. 
With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
The \texttt{Communication} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{startApp()} method in the \texttt{MainSpot} class. 
Within the \texttt{MainSpot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
\todo{Grafik zweitteilen - hier package sensornode}

\autoref{fig:UML} describes the classes of the \texttt{basestation} package.
The package \texttt{basestation} runs on the host computer and operates the basestation.
It consists of the classes \texttt{Database\-Handler} and \texttt{MainBase}.
The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
The entry point of the program is the \texttt{run()} method in the \texttt{MainBase} class. The \texttt{MainBase} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
\todo{Grafik zweitteilen - hier package basestation}

\begin{figure}[h!]
    \centering
    \includegraphics{figures/uml.pdf}
    \caption{class diagram describing the structure of the implemented SHM system}
    \label{fig:UML}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.3]{figures/teststructure.jpg}
    \caption{Laboratory setup}
    \label{fig:teststructure}
\end{figure}

This excitation method ensures a free vibration in natural frequency with little interferences.
After excitation, the sensor nodes automatically start the acceleration with a sampling rate of 40\,Hz.
To minimize the wireless data communication, each node performs a fast Fourier transform algorithm once sufficient acceleration measurements have been collected.
%Generally FFTs are used to convert equidistant samples of a function into a combination of frequencies, and the corresponding magnitudes, that has the same sample values. 
The acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the building.
Each sensor node transfers only the values of the predominating frequency---i.e. the frequency with the biggest magnitude--- \colorbox{cyan}{to the basestation}. 
The values are then stored in the MySQL database at the host computer.

@@

%\textbf{2. paragraph: data measurement and processing}
%
%\emph{\begin{itemize}
%\item learning phase
%\item excitation
%\item sensors start measuring
%\item sensor nodes \textbf{perform fft} - what is fft
%\item sensor nodes exchange frequencies
%\item neural networks check integrity
%\item if no faults detected, sensor nodes send data to basestation
%\end{itemize}}
%
%\textbf{3. paragraph: sensor fault detection}
%
%\begin{itemize}
%\item simulation of sensor fault
%\item neural networks check integrity -> error!
%\item sensor node sends alert to basestation
%\end{itemize}

>>>>

The dependability and accuracy of wireless structural health monitoring systems can be affected by data loss during wireless data transmission.
Applying adequate data compression methods can significantly reduce the amount of data that has to be transmitted and lower the occurrence of data loss.
In this paper, the design and implementation of a wireless structural health monitoring system, capable of data compression is presented. 
A fast Fourier transform is used on every sensor node to analyze oscillation frequencies. 
Only the first natural frequency of the oscillation and the correlated magnitude are transmitted to the base station.
In laboratory experiments, devised to validate the proposed approach, the data traffic is reduced by xx %.

@@

\linenumbers % Schaltet Zeilennummerierung ein
\modulolinenumbers[5] % nur jede 5. Zeile

\section*{Dictionary}

\begin{tabular}{|l|l|}
\hline 
Sensorknoten (SunSPOT) & sensor node \\ 
\hline 
einzelner Messsensor (Thermometer) & sensor \\ 
\hline 
Knoten im neuronalen Netz & neuron \\ 
\hline 
eine abgeschlossene Messungreihe & test run \\ 
\hline 
gemessene Werte & sensor data \\ 
\hline 
vorhergesagte Werte & predicted data \\ 
\hline 
durch Vorhersage erwartete Werte & expected data \\ 
\hline 
einzelner Messwert & measurement \\ 
\hline 
Test & laboratory experiments\\ 
\hline
tatsächliche, nicht virtuelle Messung & actual measurement \\
\hline
Messaufbau & test setup \\
\hline
\end{tabular}

@@

\begin{figure}[ht]
    \centering
    \includegraphics[scale=0.5]{figures/eDemoboard.png}
    \caption{Layout of the Sun SPOT sensor board \cite{spotToO}}
    \label{fig:SPOT}
\end{figure}

@@

\autoref{fig:UML-sn} describes the classes of the \texttt{sensornode} package.
The package \texttt{sensornode} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, which are embedded directly into the sensor nodes.
The \texttt{AccelerationSampler} class is responsible for measuring the acceleration.
There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
Once the acceleration exceeds a threshold, the \colorbox{cyan}{second phase is entered} by increasing the sampling rate. 
The measured values are stored into an array.
The different phases are indicatad by lighting different LEDs.
The \texttt{FFT} class performs a fast Fourier transform on the measured accelerations. 
With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
The \texttt{Communication} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{startApp()} method in the \texttt{MainSpot} class. 
Within the \texttt{MainSpot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.

\begin{figure}[h!]
    \centering
    \includegraphics{figures/uml-sensornode.pdf}
    \caption{class diagram of the sensornode package}
    \label{fig:UML-sn}
\end{figure}

\autoref{fig:UML-bs} describes the classes of the \texttt{basestation} package.
The package \texttt{basestation} runs on the host computer and operates the basestation.
It consists of the classes \texttt{Database\-Handler} and \texttt{MainBase}.
The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
The entry point of the program is the \texttt{run()} method in the \texttt{MainBase} class. The \texttt{MainBase} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
\todo{Grafik zweitteilen - hier package basestation}

\begin{figure}[h!]
    \centering
    \includegraphics{figures/uml-basestation.pdf}
    \caption{class diagram of the basestation package}
    \label{fig:UML-bs}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.4]{figures/teststructure.jpg}
    \caption{Laboratory setup}
    \label{fig:teststructure}
\end{figure}

This excitation method ensures a free vibration in natural frequency with little interferences.
After excitation, the sensor nodes automatically start measuring acceleration with a sampling rate of 40\,Hz.
To minimize the wireless data communication, each node performs a fast Fourier transform algorithm once sufficient acceleration measurements have been collected.

The acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the building.
Each sensor node transfers only the values of the predominating frequency---i.e. the frequency with the biggest magnitude--- \colorbox{cyan}{to the basestation}. 
The values are then stored in the MySQL database at the host computer.

@@
